Relict deciduous oaks show high resilience to drought despite the effect of land-use legacies

### 2019-02-21 man/ms.Rmd
update

The analysis of growth changes revealed differences in forest history between sites (Figures 5 and S4). Northern site (SJ) showed two release events (GC > 50 %) detected at stand-wise scale (occurring in more than 50 % of sampled trees): the first during the 1940 decade and the second in the period 1994-2001. These periods alternate with periods of supression. Southern sites (CA-High and CA-Low) showed no release event except one for CA-High at the beginning of the `REVIEW`1840s? and no suppression event in the last 50 years.

@@

Resilience (*Rs*) values observed in our study sites for the latest drought events in both tree-growth and greenness were close to 1, except for the tree-growth during 2005 event (Tables S1 and S2). Despite the 2012 drought event was more severe and intense than 2005, according with our analysis with SPEI data (Table S3), resilience values for greenness and tree-growth were greater for 2012 than for 2005. These results could be explained by the different timing of the two droughts, *i.e.* the 2012 drought was a winter drought [@Trigo2013] that occurred earlier than the 2005 drought. The latter lasted less than 2012 drought, but matched the period of maximum growth for *Q. pyrenaica* forests [@PeresLis2017]. Our results highlight the importance of the drought timing on the tree-growth responses, as have been indicated for other Mediterranean-tree species [*e.g.* *Q. ilex*, @Camarero2015b].

@@

>>>>

The analysis of growth changes revealed differences in forest history between sites (Figures 5 and S4). Northern site (SJ) showed two release events (GC > 50 %) detected at stand-wise scale (occurring in more than 50 % of sampled trees): the first during the 1940 decade and the second in the period 1995-2000. These periods alternate with periods of supression. Southern sites (CA-High and CA-Low) showed no release events except for CA-High at the beginning of the 1830 and no suppression event in the last 50 years.

@@

Resilience (*Rs*) values observed in our study sites for the latest drought events in both tree-growth and greenness were close to 1, except for the tree-growth during 2005 event (Tables S1 and S2). Despite the 2012 drought event was more severe and intense than 2005, according with our analysis with SPEI data (Table S3), resilience values for greenness and tree-growth were greater for 2012 than for 2005. These results could be explained by the different timing of the two droughts, *i.e.* the 2012 drought was a winter drought [@Trigo2013] that occurred earlier than the 2005 drought. The latter lasted less than 2012 drought, but matched the period of maximum growth for *Q. pyrenaica* forests [@PeresLis2017]. Our results are in line with recently studies that indicate the timing of the drought as a key factor determining tree recovery after drought [@Huang2018],  that also has been observed for other Mediterranean oak species [*e.g.* Holm oak; @Camarero2015b].

@@

--> Ver y resumir páginas 18-19 Notas Naturkunde
